---
title: "Advanced Data Visualization"
author: "James Cantonwine"
date: "7/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## plotly

### Interactive plots

`plotly` is a package for translating `ggplot2` graphs into interactive or custom web-based visualizations. Once a plot has been created and edited within `gplot`, using the function `ggplotly()` will create an interactive plot:
```{r}
## Load libraries
library(ggplot2)
library(plotly)

## Generate ggplot object
p <- ggplot(iris, aes(x = factor(Species), y = Petal.Length, color = Petal.Width)) +
  geom_boxplot() +
  geom_point()

## Generate interactive graphic
p <- ggplotly(p)
p
```

The [plotly docmentation](https://plotly-r.com/) will be a helpful resource.

### Animated plots

Plots can be animated by specifying the `frame` argument. This example uses data from the `gapminder` package. Because the frame argument is set to `year`, the plot will cycle through years available within the data set.
```{r, message=FALSE}
## install the package 
install.packages("gapminder")

## get the data
data(gapminder, package = "gapminder")

## generate ggplot object
gg <- ggplot(gapminder, aes(gdpPercap, lifeExp, color = continent)) +
  geom_point(aes(frame = year, ids = country)) +
  scale_x_log10()

## plot animated graphic
ggplotly(gg)
```

Objects created with `plotly` can be printed into .Rmd as long as the document knits to HTML. If a `plotly` is knit into a non-HTML format, it will print as a static .png screenshot.

## gganimate

There are other packages that can generate animated graphics. One advantage to using `gganimate` is that the output can be direct to a .gif, .mp4, .swf, or .html object. 

**The version of gganimate used here refers to an older API and will not work properly: either the version of gganimate needs to be downgraded or the code updated.**
```{r, message=FALSE}
## download and load gganimate 
devtools::install_github("dgrtwo/gganimate")
library(gganimate)

## plot animated graphic using gganimate
gganimate(gg)

## to save this plot
gganimate(gg, "output.gif")
```
