---
title: "Multiple Plots in R"
author: "James Cantonwine"
date: "7/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## Installing patchwork

These examples will use the `patchwork` package. Because `patchwork` is not on the CRAN database, installation follows a different procedure:
```{r, message=FALSE}
install.packages("devtools")

library(devtools)
devtools::install_github("thomasp85/patchwork")
```

More information about `patchwork` can be found on the [Patchwork GitHub](https://github.com/thomasp85/patchwork).

## Basic plotting using patchwork

After creating two plots, `patchwork` allows for these to be combined using a `+` operator:
```{r, message=FALSE}
library(ggplot2)
library(patchwork)

p1 <- ggplot(mtcars) + geom_point(aes(wt, mpg))
p2 <- ggplot(mtcars) + geom_boxplot(aes(gear, mpg, group = gear))

p1 + p2
```

## Altering the layout

The `plot_layout()` function allows for the specification of the number of columns or rows with `ncol = ` and `nrow = ` respectively:
```{r, message=FALSE}
p1 + p2 + plot_layout(ncol = 1)
```

The `heights` argument allows for control of the height of each plot. In the example below, the first plot will be 3 times the height of the second:
```{r, message=FALSE}
p1 + p2 + plot_layout(ncol = 1, heights = c(3, 1))
```

Nested plots can also be created using `patchwork` through brackets:
```{r, message=FALSE}
p3 <- ggplot(mtcars) + geom_smooth(aes(disp, qsec))
p4 <- ggplot(mtcars) + geom_bar(aes(carb))

p4 + (
  p1 + (
    p2 +
    p3 +
    plot_layout(ncol = 1)
  )
) +
  plot_layout(ncol = 1)
```

## Additional operators

Using `|` will always add plots together from left to right, while `/` will stack plots on top of one another:
```{r, message=FALSE}
(p1 | p2 | p3) / p4
```

Using `&` will apply a theme to all plots, while `*` will only apply the theme to plots in that nesting level:
```{r, message=FALSE}
p1 + (p2 + p3) + p4 + plot_layout(ncol = 1) & theme_bw()

(p1 + (p2 + p3) + p4 + plot_layout(ncol = 1)) * theme_bw()
```