---
title: "Tables in R"
author: "James Cantonwine"
date: "7/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)
```

## Exploratory tables

This code establishes a new dataframe `df` from the `diamonds` dataset. `df` groups the data by the quality of the cut (`cut`) and uses a `dplyr` function to summarize the data. The summary will include the number of diamonds in each group as well as the minimum, mean, and maximum prices within each group.

```{r}
df <- diamonds %>% 
  group_by(cut) %>%
  dplyr::summarize(
    N = n(), 
    min = min(price), 
    avg = mean(price), 
    max = max(price)
  )
df
```

Good qualities of this table include:
* A reasonable number of rows and columns for understanding
* The first column `cut` is organized logically from lowest to highest quality
* Comparisons within the data are made vertically

A few changes would improve this table from exploratory to explanatory:
* The column headers could be more clear
* The table should have a title or caption
* It could be more aesthetically pleasing

## Improving table outputs

By default R uses a monospaced font for displaying outputs. Using the `kable()` function from the `knitr` package improves readability in both the console and in Markdown.
```{r}
library(knitr)
kable(df)
```

The `add_header_above` function within the `kableExtra` package can specify that the three price columns all refer to US dollars. We can also  use `digits` to specify the number of significant digits we want the output to display. To further improve readability, we will use `kable_styling()` to shade alternate rows.
```{r}
library(kableExtra)
kable(df, digits = 0, "html") %>%
  kable_styling("striped", "bordered") %>%
  add_header_above(c(" " = 2, "price (USD)" = 3))
```

## Annotating a table

The `kable` package allows for a `caption` argument. `kableExtra` also has a `footnote()` function that can be used to include text at the bottom of the table - generally the source of the data.
```{r}
kable(df, digits=0, "html", caption="Table 1: Diamonds Price by Quality of Cut. Most\
 Diamonds are of the highest quality cut and the most expensive diamonds are of the \
highest quality") %>%
  kable_styling("striped", "bordered") %>% 
  add_header_above(c(" " = 2,  "price (USD)" = 3)) %>% 
  footnote(general="Diamonds dataset from ggplot2", general_title="Source:",footnote_as_chunk = T)
```

