---
title: "Saving Plots"
author: "James Cantonwine"
date: "7/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## Image Types

* JPEG - takes up less space, will lose quality with each save, have a white background by default
* PNG - takes up more space, preserves quality, has a transparent background by default, may show pixels when magnified
* TIFF - takes up more space, preserves quality, can be edited easily
* PDF - works well with multiple images per file, prints well and consistently, can scale without pixelation

## Saving with ggsave

The `ggsave()` function allows for files to be saved in mutiple file types and in specified locations as below:
```{r}
## generate plot
myplot <- ggplot(data = mtcars) + geom_point(mapping = aes(x = wt, y = mpg))

## save plot as PNG
ggsave(plot = myplot, filename = "myplot.png", path = "figures/exploratory_figures")

## save plot as JPEG
ggsave(plot = myplot, filename = "myplot.jpeg", path = "figures/exploratory_figures")
```

The `ggsave()` function can also be used to specify image dimensions. This code will overwrite the prior PNG image and save a version with an increased size:
```{r}
## save plot as larger PNG
ggsave(plot = myplot, filename = "myplot.png", path = "figures/exploratory_figures", height = 9, width = 9, unit = c("in"))
```

## Alternatives

Specific functions to save exist for different file types and can be used if `ggsave()` presents any issues. Thes include `png()`, `jpeg()`, `pdf()`, and `tiff()`. Specify the save path within the function, use `print`, and then end with `dev.off()` as shown below:
```{r}
jpeg("figures/exploratory_figures/myplot.jpeg")
print(myplot)
dev.off()
```

