---
title: "Data Visualization - Basics of ggplot2"
author: "James Cantonwine"
date: "7/19/2019"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## General form of plots with ggplot2:
ggplot(data = DATASET) + 
  geom_PLOT_TYPE(mapping = aes(VARIABLE(S)))

## Scatterplots with geom_point()

### A basic plot with the diamonds dataset
```{r}
## The function geom_point plots points for a scatterplot
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = carat, y = price))
```

### Mapping point color to a variable
```{r}
## Putting color as a variable within aes maps it to a variable
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = carat, y = price, color = clarity))
```

### Manually specifying a point color
```{r}
## Here the color is specified outside of the aesthetic function and applied to all points
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = carat, y = price), color = "red")
```

We can also use **size** and **shape** as specifications in the same way - inside aes() when mapped to a variable and outside it when applied to the entire plot. There are 25 different shapes available for specification.

## Using facets to categorize data
```{r}
## The tilde is read as "by" - that variable used to group the data into facets
ggplot(data = diamonds) + 
  geom_point(mapping = aes(x = carat, y = price)) + 
  facet_wrap(~clarity, nrow = 2)
```

## Additional geom_ functions

### geom_histogram
```{r}
ggplot(data = diamonds) + 
  geom_histogram(mapping =  aes(carat))
```

This can be used to help verify assumptions about frequency, as points on a scatterplot may stack or overlap.

### geom_bar
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(cut))
```

This allows for comparison of a set of numbers and a categorical variable.

### geom_boxplot
```{r}
ggplot(data = diamonds) + 
  geom_boxplot(aes(x = color, y = price))
```

